---
title: "ECON 4198W Merge and Analysis"
format: html
editor: visual
---

## Merging the two datasets

```{r}
library(data.table)

CBP_data <- as.data.table(read.csv(("cbp_2012_2022.csv")))
H1b_2016 <- as.data.table(read.csv("Lottery_Applications_vs_Approvals.csv"))
```

### Cleaning

```{r}
H1b_2016 <- H1b_2016[H1b_2016$num_petitions != 0]

library(dplyr)

H1b_2016 <- H1b_2016 |>
  mutate(win_rate = initial_approval / num_petitions)
```

### Merge

```{r}
ProsemData_Raw <- CBP_data |>
  mutate(COUNTY = substr(GEO_ID, 10, 14)) |>
  left_join(H1b_2016, by = "COUNTY") |>
  mutate(h1b_activity = ifelse(is.na(win_rate), 0, 1))

ProsemData_Raw <- ProsemData_Raw[ProsemData_Raw$num_petitions >= 20]

ProsemData_Raw <- ProsemData_Raw |>
  distinct()
```

## Fixing Duplicates in County Business Patterns

```{r}
check_panel <- function(df, unit = "GEO_ID", time = "YEAR", sector = "NAICS") {
  library(dplyr)
  dupes <- df %>%
    group_by(across(all_of(c(unit, time, sector)))) %>%
    filter(n() > 1)
  
  if (nrow(dupes) == 0) {
    message("✅ Data is uniquely identified by unit, time, and sector.")
  } else {
    message("❌ Found duplicate rows:")
    print(dupes)
  }
}

duplicates <-check_panel(ProsemData_Raw)
```

```{r}
ProsemData <- ProsemData_Raw |>
  group_by(GEO_ID, YEAR, NAICS) |>
  summarise(
    across(c(NAME, COUNTY, initial_approval, num_petitions, win_rate, h1b_activity), first),  # keep first occurrence
    ESTAB = sum(ESTAB, na.rm = TRUE),
    .groups = "drop"
  )

check_panel(ProsemData)

#write_dta("ProsemData.dta", data = ProsemData)
```

```{r}
ProsemData <- ProsemData |>
  mutate(Post = ifelse(YEAR >= 2017, 1, 0))
```

Merge and remove counties with number of petitions less than 20 to reduce noise in the data.

## Creating Diversity Measures

```{r}
library(tidyr)

calculate_indices <- function(ProsemData){
  total <- sum(ProsemData$ESTAB)
  p_i <- ProsemData$ESTAB / total
  
  # Shannon Index: -sum(p_i * ln(p_i))
  shannon <- -sum(p_i * log(p_i))
  
  # Herfindahl-Hirschman Index: sum(p_i^2)
  hhi <- sum(p_i^2)*10000
  
  return(data.frame(shannon_index = shannon, hhi = hhi, ESTAB = total))
}

H1bdf <- ProsemData |>
  group_by(COUNTY, NAME, YEAR, win_rate) |>
  group_modify(~ calculate_indices(.x)) |>
  ungroup()
```

## Last Minute Data Cleaning + Adding State Column

```{r}
H1bdf <- H1bdf |>
  mutate(win_rate = ifelse(win_rate > 1, 1, win_rate))

# Create a lookup table for state FIPS codes and abbreviations
state_fips_lookup <- data.frame(
  state_fips = sprintf("%02d", c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56)),
  state_abbr = c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")
)

# Extract the state FIPS code from the county FIPS code and join with the lookup table
H1bdf <- H1bdf %>%
  mutate(state_fips = substr(COUNTY, 1, 2)) %>%
  left_join(state_fips_lookup, by = "state_fips") %>%
  rename(STATE = state_abbr) %>%
  select(-state_fips)

library(haven)
#write.csv(H1bdf, "h1bdf.csv", row.names = FALSE)
#write_dta("H1bdf.dta", data = H1bdf)

```

### Extra last minute adding time lags

```{r}
H1bdf <- H1bdf |>
  group_by(COUNTY) |>
  arrange(YEAR) |>
  mutate(hhi_lagy1 = lag(hhi, 1)) |>
  mutate(hhi_lagy2 = lag(hhi, 2)) |>
  ungroup()
```

```{r}
CBP_data_2022 <- CBP_data |>
  filter(YEAR == 2022)

calculate_indices <- function(CBP_data_2022){
  total <- sum(CBP_data_2022$ESTAB)
  p_i <- CBP_data_2022$ESTAB / total
  
  # Shannon Index: -sum(p_i * ln(p_i))
  shannon <- -sum(p_i * log(p_i))
  
  # Herfindahl-Hirschman Index: sum(p_i^2)
  hhi <- sum(p_i^2)*10000
  
  return(data.frame(shannon_index = shannon, hhi = hhi, ESTAB = total))
}

CBP_diversity <- CBP_data_2022 |>
  group_by(GEO_ID, NAME, YEAR) |>
  group_modify(~ calculate_indices(.x)) |>
  ungroup()

CBP_diversity <- CBP_diversity |>
  mutate(
    log_hhi = log(hhi),
    log_shannon = log(shannon_index)
  )
```

## Descriptive Statistics

```{r}
summary(H1bdf[c("shannon_index", "hhi", "win_rate")])

H1bdf |> 
  group_by(YEAR) |>
  summarise(
    mean_shannon = mean(shannon_index, na.rm=TRUE),
    sd_shannon = sd(shannon_index),
    mean_hhi = mean(hhi, na.rm = TRUE),
    mean_win_rate = mean(win_rate, na.rm = TRUE)
  )
```

```{r}
diversityrank <- H1bdf |>
  filter(YEAR == 2022) |>
  mutate(HHI_rank = min_rank(hhi))  # negative sign reverses rank

sd(diversityrank$hhi, na.rm = TRUE)
sd(diversityrank$hhi, na.rm = TRUE)/4
```

```{r}
diversityrank_full <- CBP_diversity |>
  mutate(HHI_rank = min_rank(hhi))

```

### Visualizations

```{r}
library(ggplot2)

shannon_dist <- ggplot(H1bdf, aes(x = shannon_index)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Shannon Diversity Index", x = "Shannon Index", y = "Frequency")

hhi_dist <- ggplot(H1bdf, aes(x = hhi)) +
  geom_histogram(bins = 30, fill = "salmon", color = "black") +
  labs(title = "Distribution of HHI", x = "HHI", y = "Count")

winrate_dist <- ggplot(H1bdf, aes(x = win_rate)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  labs(x = "Win Rate", y = "Count") +
  theme_classic()

shannon_dist
hhi_dist
winrate_dist

ggsave("Figures/shannon_distribution.pdf", shannon_dist, width = 7, height = 5)
ggsave("Figures/hhi_distribution.pdf", hhi_dist, width = 7, height = 5)
ggsave("Figures/winrate_distribution.pdf", winrate_dist, width = 7, height = 5)

```

Here are the distributions for the diversity measures. Clustering for Shannon is very high and while most HHIs are clustering around the same, there are some large outliers. This may require applying a natural log on the diversity measures.

## Regressions

```{r}
H1bdf <- H1bdf |>
  mutate(
    log_shannon = log(shannon_index),
    log_hhi = log(hhi),
    log_estab = log(ESTAB)
  )

summary(H1bdf[c("win_rate", "log_shannon", "log_hhi")])
```

Before doing anything, I'm gonna log my diversity measures because the interpretation makes less sense if I don't.

### New Visualizations

```{r}
log_shannon_dist <- ggplot(H1bdf, aes(x = log_shannon)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Distribution of log(Shannon Index)")

log_hhi_dist <- ggplot(H1bdf, aes(x = log_hhi)) +
  geom_histogram(bins = 30, fill = "red", color = "black") +
  labs(title = "Distribution of log(HHI)")

log_shannon_dist
log_hhi_dist

ggsave("Figures/log_shannon_distribution.pdf", log_shannon_dist, width = 7, height = 5)
ggsave("Figures/log_hhi_distribution.pdf", log_hhi_dist, width = 7, height = 5)
```

### Preliminary Regressions

```{r}
library(lmtest)
library(sandwich)

reg_hhi <- lm(log_hhi ~ win_rate, data = H1bdf)
reg_shannon <- lm(log_shannon ~ win_rate, data = H1bdf)

summary(reg_hhi)
summary(reg_shannon)

coeftest(reg_hhi, vcov = vcovHC(reg_hhi, type = "HC1"))
coeftest(reg_shannon, vcov = vcovHC(reg_shannon, type = "HC1"))
```

Regressions show high P values but low $R^2$, meaning there is likely a lot of omitted variable bias. Next is to control for county and year fixed effects.

```{r}
library(fixest)

reg_hhifixed <- feols(log_hhi ~ win_rate | YEAR, data = H1bdf, cluster = ~COUNTY)

lm_hhifixed <- lm(log_hhi ~ win_rate + factor(COUNTY) + factor(YEAR), data = H1bdf)

reg_shannonfixed <- feols(log_shannon ~ win_rate | + YEAR, data = H1bdf, cluster = ~COUNTY)

lm_shannonfixed <- lm(log_shannon ~ win_rate + factor(COUNTY) + factor(YEAR), data = H1bdf)

summary(reg_hhifixed)
summary(reg_shannonfixed)

#summary(lm_hhifixed)
#summary(lm_shannonfixed)

#coeftest(reg_hhifixed, vcov = vcovHC(reg_hhifixed, type = "HC1"))
#coeftest(reg_shannonfixed, vcov = vcovHC(reg_shannonfixed, type = "HC1"))
```

After controlling for county and year fixed effects, the $R^2$ jumped significantly to 0.85 for HHI and 0.79 for Shannon, which means a lot of omitted variable bias has been controlled for with county and year fixed effects.

## Differences in Differences

### Parallel Trends

#### Creating Treatment and Post Variables

```{r}
H1bdf <- H1bdf %>%
  mutate(Post = ifelse(YEAR >= 2017, 1, 0))  # Post-treatment period

```

#### Creating Relative Year Variable

```{r}
H1bdf <- H1bdf %>%
  mutate(
    event_time = YEAR - 2017,  # Define event time relative to treatment year
    event_time = ifelse(event_time < -5, -5, event_time),  # Group <= -5 into one category
    event_time = ifelse(event_time > 5, 5, event_time)  # Group >= 5 into one category
  )
```

## Event Study Regression

```{r}
event_study_hhi <- feols(log_hhi ~ i(event_time, win_rate, ref = -1) + factor(COUNTY) + factor(YEAR), data = H1bdf, cluster = "COUNTY")

event_study_shannon <- feols(log_shannon ~ i(event_time, win_rate, ref = -1) + factor(COUNTY) + factor(YEAR), data = H1bdf, cluster = "COUNTY")

wald(event_study_hhi, "event_time::[-5,-4,-3,-2, -1]")

summary(event_study_hhi)
summary(event_study_shannon)

# iplot() draws on the base graphics device, so wrap it to draw it twice:
# once for the knitted output and once into the PDF written for the paper.
plot_event_study_hhi <- function() {
  # main = "" rather than dropping the argument: iplot's default title is
  # "Effect on log_hhi", so omitting it leaves a title rather than removing one.
  # Shrinking the top margin reclaims the band the title used to occupy.
  op <- par(mar = c(4.5, 4.5, 1, 1))
  on.exit(par(op))

  iplot(event_study_hhi,
        main = "",
        xlab = "Years Relative to Event",
        ylab = "Effect of Win Rate on Log(HHI)")
}

plot_event_study_hhi()

pdf("Figures/event_study_hhi.pdf", width = 7, height = 5)
plot_event_study_hhi()
dev.off()
```

## Event Study Business Counts

```{r}
event_study_counts <- feols(log_estab ~ i(event_time, win_rate, ref = -1) | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

summary(event_study_counts)

wald(event_study_counts, "event_time::[-5,-4,-3,-2, -1]")

plot_event_study_counts <- function() {
  # NB: `ci` is not an iplot argument - it lands in `...` and fixest reads it as
  # a second model, which errors. The confidence level is `ci_level`; the
  # default 0.95 is what the paper's caption describes.
  op <- par(mar = c(4.5, 4.5, 1, 1))
  on.exit(par(op))

  iplot(event_study_counts,
        main = "",
        xlab = "Years Relative to Event",
        ylab = "Effect of Win Rate on Log(Counts)")
}

plot_event_study_counts()

pdf("Figures/event_study_counts.pdf", width = 7, height = 5)
plot_event_study_counts()
dev.off()
```

Parallel trends assumption holds because pre-treatment coefficients are all near-zero and statistically insignificant. Post-treatment effects show a causal impact of an H-1B visa increasing business diversity.

## Continuous Difference in Differences

```{r}
# Set panel structure (COUNTY = panel ID, YEAR = time variable)
H1bdf <- H1bdf %>%
  mutate(COUNTY = as.factor(COUNTY), YEAR = as.integer(YEAR))

did_continuous_hhi <- feols(log_hhi ~ win_rate * Post | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

summary(did_continuous_hhi)
```

The P value is near 0 using, meaning that this coefficient is highly significant. This means that a 1 percentage point increase in H-1B win rate is associated with a 0.063% decline in the HHI index. Or, an increase in H-1B win rates increases business diversity.

Standard Error Treament used was Driscoll-Kraay, which is robust to spatial and temporal dependence.

## Sensitivity: Cap-Exempt Filter Breadth

How much of the result depends on where the cap-exempt line is drawn. The three specs are built in `H-1B_Data_Cleaning.Rmd`: no exclusion at all, the published keyword list, and a deliberately over-inclusive list. Only the *dose* changes across specs — the HHI outcome is built from a county's own NAICS establishment shares and never touches the filter — so the county-year panel is built once and the dose swapped in.

```{r}
library(fixest)
library(broom)

specs <- c("none", "baseline", "broad")

# colClasses is not optional here. COUNTY is a zero-padded FIPS string, and
# read.csv would turn "01001" into the integer 1001, silently failing to merge
# every leading-zero county. The existing read at the top of this file survives
# only because a stray "Unknown" value forces the column to character -- if that
# value ever disappears the join breaks with no error, just missing counties.
dose_raw <- read.csv("Lottery_Applications_vs_Approvals.csv",
                     colClasses = c(COUNTY = "character"))

# Outcome panel: independent of the cap-exempt filter, so built once.
cbp_panel <- CBP_data |>
  mutate(COUNTY = substr(GEO_ID, 10, 14)) |>
  distinct() |>
  group_by(GEO_ID, YEAR, NAICS) |>
  summarise(across(c(NAME, COUNTY), first),
            ESTAB = sum(ESTAB, na.rm = TRUE), .groups = "drop") |>
  group_by(COUNTY, NAME, YEAR) |>
  group_modify(~ calculate_indices(.x)) |>
  ungroup() |>
  mutate(log_hhi = log(hhi),
         log_shannon = log(shannon_index),
         log_estab = log(ESTAB))
```

```{r}
# inner_join makes explicit what the main pipeline does by accident: counties with
# no H-1B activity have NA petitions, and `NA >= 20` is NA, so they are dropped by
# the threshold rather than by any deliberate decision.
build_panel <- function(spec, min_petitions = 20, keep_counties = NULL) {
  dose <- dose_raw |>
    transmute(
      COUNTY,
      initial_approval = .data[[paste0("initial_approval_", spec)]],
      num_petitions    = .data[[paste0("num_petitions_", spec)]]
    ) |>
    filter(!is.na(num_petitions), num_petitions >= min_petitions) |>
    mutate(win_rate = pmin(initial_approval / num_petitions, 1))

  if (!is.null(keep_counties)) dose <- filter(dose, COUNTY %in% keep_counties)

  cbp_panel |>
    inner_join(dose, by = "COUNTY") |>
    mutate(Post = as.integer(YEAR >= 2017),
           event_time = pmin(pmax(YEAR - 2017, -5), 5))
}

# Each spec admits a different set of counties (a broader filter shrinks the
# denominator and pushes counties under the 20-petition floor). Reporting both the
# spec's own sample and the sample common to all three separates "the dose changed"
# from "the sample changed".
common_counties <- Reduce(intersect, lapply(specs, function(s) unique(build_panel(s)$COUNTY)))
length(common_counties)
```

```{r}
run_one <- function(spec, sample_label, keep_counties = NULL) {
  d <- build_panel(spec, keep_counties = keep_counties)
  m <- feols(log_hhi ~ win_rate * Post | COUNTY + YEAR, data = d, cluster = "COUNTY")
  tt <- tidy(m) |> filter(term == "win_rate:Post")

  tibble(
    spec          = spec,
    sample        = sample_label,
    counties      = n_distinct(d$COUNTY),
    med_petitions = median(d$num_petitions),
    mean_win_rate = mean(d$win_rate),
    estimate      = tt$estimate,
    std.error     = tt$std.error,
    p.value       = tt$p.value
  )
}

sens_table <- bind_rows(
  lapply(specs, run_one, sample_label = "own"),
  lapply(specs, run_one, sample_label = "common", keep_counties = common_counties)
)

sens_table

# The baseline spec must reproduce the published estimate exactly; if it does not,
# the refactor above has drifted from the main pipeline and nothing else here is
# trustworthy.
all.equal(
  sens_table$estimate[sens_table$spec == "baseline" & sens_table$sample == "own"],
  unname(coef(did_continuous_hhi)["win_rate:Post"]),
  tolerance = 1e-8
)
```

```{r}
# Event study under each spec, overlaid. ggplot rather than iplot() because three
# base-graphics event studies cannot be drawn on one set of axes.
es_tidy <- bind_rows(lapply(specs, function(s) {
  m <- feols(log_hhi ~ i(event_time, win_rate, ref = -1) | COUNTY + YEAR,
             data = build_panel(s), cluster = "COUNTY")
  tidy(m, conf.int = TRUE) |>
    filter(grepl("^event_time::", term)) |>
    mutate(event_time = as.numeric(sub("^event_time::(-?[0-9]+):win_rate$", "\\1", term)),
           spec = s)
}))

# Reinstate the omitted reference period so the plotted path is not misread.
es_tidy <- bind_rows(
  es_tidy,
  tibble(event_time = -1, estimate = 0, conf.low = 0, conf.high = 0, spec = specs)
)

es_plot <- ggplot(es_tidy, aes(event_time, estimate, colour = spec, shape = spec)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.3) +
  geom_vline(xintercept = -0.5, linetype = "dotted", linewidth = 0.3) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  position = position_dodge(width = 0.45)) +
  scale_x_continuous(breaks = -5:5) +
  labs(x = "Years Relative to Event",
       y = "Effect of Win Rate on Log(HHI)",
       colour = "Cap-exempt filter", shape = "Cap-exempt filter") +
  theme_minimal()

es_plot

ggsave("Figures/sensitivity_cap_exempt.pdf", es_plot, width = 7, height = 5)
```

## Heterogeneity

### Mediation Regressions

```{r}
library(modelsummary)

feols(log_hhi ~ win_rate * Post | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

feols(log_estab ~ win_rate * Post | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

feols(log_hhi ~ win_rate * Post + log_estab | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

models <- list(
  "log(ESTAB)" = feols(log_hhi ~ win_rate * Post | COUNTY + YEAR, cluster = "COUNTY", data = H1bdf),
  "log(HHI)" = feols(log_estab ~ win_rate * Post | COUNTY + YEAR, cluster = "COUNTY", data = H1bdf),
  "log(HHI) w/ mediator" = feols(log_hhi ~ win_rate + log_estab | COUNTY + YEAR, cluster = "COUNTY", data = H1bdf)
)

modelsummary(models,
             stars = TRUE,
             coef_rename = c("win_rate" = "Win Rate", "log(ESTAB)" = "log(Counts)"),
             gof_omit = "Adj|AIC|BIC|Log.Lik|RMSE",
             output = "latex",
             notes = "Standard errors clustered by county in parentheses. *** p<0.01, ** p<0.05, * p<0.10")
```

### Firm Level Data + NAICS Fixed Effects

```{r}
feols(log(ESTAB) ~ win_rate * Post | COUNTY + YEAR + NAICS, cluster = "COUNTY", data = ProsemData)
```

### Immigrant heavy industries

```{r}
abs_data <- as.data.table(read.csv("Annual_Business_Survey.csv"))

abs_data <- abs_data[OWNCHAR == "EM"]

naics_counts <- abs_data[, .N, by = NAICS2022][order(-N)]
```

```{r}
setDT(ProsemData)

ImmigrantHeavy <- ProsemData[, skill_group := fcase(
  NAICS %in% c(6211, 5412, 5413, 5415, 5416, 5112, 5182, 5191, 5242), "High",
  NAICS %in% c(5617, 5613, 7225, 7211, 8121, 2383, 2382, 2389, 6216, 6231, 6232, 6233, 6241, 6244, 4853, 3118, 4461, 4451, 4481, 4471, 8111, 8123), "Low",
  default = "Other"
)]

ImmigrantHeavy <- ImmigrantHeavy[skill_group %in% c("Low", "High")]
LowSkill <- ImmigrantHeavy[skill_group == "Low"]
HighSkill <- ImmigrantHeavy[skill_group == "High"]

```

```{r}
did_immigrant <- feols(log(ESTAB) ~ win_rate * Post | COUNTY + YEAR + NAICS, data = ImmigrantHeavy, cluster = "COUNTY")

did_skilled <- feols(log(ESTAB) ~ win_rate * Post | COUNTY + YEAR + NAICS, data = HighSkill, cluster = "COUNTY")

did_unskilled <- feols(log(ESTAB) ~ win_rate * Post | COUNTY + YEAR + NAICS, data = LowSkill, cluster = "COUNTY")

summary(did_immigrant)
summary(did_skilled)
summary(did_unskilled)
```

```{r}
library(broom)

# Extract coefficient and SE from each model
immigrant_coef <- tidy(did_immigrant) |> filter(term == "win_rate:Post")
highskill_coef <- tidy(did_skilled) |> filter(term == "win_rate:Post")
lowskill_coef <- tidy(did_unskilled) |> filter(term == "win_rate:Post")

# Combine into a single dataframe
heterogeneity_df <- rbind(
  data.frame(Group = "Immigrant-heavy",
             Estimate = immigrant_coef$estimate,
             SE = immigrant_coef$std.error),
  data.frame(Group = "High-skill",
             Estimate = highskill_coef$estimate,
             SE = highskill_coef$std.error),
  data.frame(Group = "Low-skill",
             Estimate = lowskill_coef$estimate,
             SE = lowskill_coef$std.error)
)
```

#### Heterogeneity Visualization

```{r}
library(ggplot2)

ggplot(heterogeneity_df, aes(x = Group, y = Estimate)) +
  geom_col(fill = "steelblue") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE),
                width = 0.2, color = "black") +
  geom_text(aes(label = sprintf("%.3f", Estimate)),
            vjust = -0.5, size = 4) +
  labs(title = "H-1B Lottery Effects Heterogeneity",
       y = "Effect on log(Business Counts)",
       x = "") +
  theme_classic()
```

## Robustness Checks

### Placebo Test

```{r}
H1bdf <- H1bdf %>%
  mutate(Post_Placebo = ifelse(YEAR >= 2014, 1, 0))

placebo_model <- feols(log_hhi ~ win_rate * Post_Placebo + i(COUNTY) + i(YEAR), data = H1bdf, cluster = "COUNTY")

summary(placebo_model)
```

P value of placebo test is not statistically significant at the 10% level.

### Interpretation

```{r}
sd_logshannon <- sd(H1bdf$log_shannon, na.rm = TRUE)
effect_shannon <- 0.0265  # or whatever your DiD coefficient is
percent_sd <- (effect_shannon / sd_logshannon) * 100

sd_loghhi <- sd(H1bdf$log_hhi, na.rm = TRUE)
effect_hhi <- -0.0657
percent_sd_hhi <- (abs(effect_hhi) / sd_loghhi) * 100

print(percent_sd_hhi)
print(percent_sd)
```

### NAICS Exclusion

```{r}
ProsemData <- ProsemData |>
  mutate(
    NAICS2 = substr(NAICS, 1, 2)
  )

# Define sets of 2-digit NAICS codes
local_services <- c("44","45","62","71","72","81")
tradables <- c("31","32","33","42","51","52","54")

ProsemData_Local <- ProsemData |>
  filter(NAICS2 %in% local_services)

ProsemData_Spillover <- ProsemData |>
  filter(NAICS2 %in% tradables)

H1bdf_local <- ProsemData_Local |>
  group_by(COUNTY, NAME, YEAR, win_rate) |>
  group_modify(~ calculate_indices(.x)) |>
  ungroup()

H1bdf_spillover <- ProsemData_Spillover |>
  group_by(COUNTY, NAME, YEAR, win_rate) |>
  group_modify(~calculate_indices(.x)) |>
  ungroup()

H1bdf_local <- H1bdf_local |>
  mutate(Post = ifelse(YEAR >= 2017, 1, 0))

H1bdf_spillover <- H1bdf_spillover |>
  mutate(Post = ifelse(YEAR >= 2017, 1, 0))
```

```{r}
feols(log(hhi) ~ win_rate*Post | COUNTY + YEAR, data = H1bdf_local, cluster = "COUNTY")

feols(log(hhi)~ win_rate*Post | COUNTY + YEAR, data = H1bdf_spillover, cluster = "COUNTY")
```

### DiD with Lags

```{r}
feols(log(hhi) ~ win_rate*Post + log(hhi_lagy1) + log(hhi_lagy2) | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

```


```{r}
library(broom)

# Example: store your models
m1 <- feols(log(hhi) ~ win_rate*Post | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")
m2 <- feols(log(hhi) ~ win_rate*Post + log(hhi_lagy1) + log(hhi_lagy2) | COUNTY + YEAR,
            data = H1bdf, cluster = "COUNTY")

# Extract tidy results with confidence intervals
tidy_m1 <- broom::tidy(m1, conf.int = TRUE) %>%
  mutate(model = "No Lags")
tidy_m2 <- broom::tidy(m2, conf.int = TRUE) %>%
  mutate(model = "With Lags")

# Combine
tidy_all <- bind_rows(tidy_m1, tidy_m2) %>%
  filter(term %in% c("win_rate:Post", "log(hhi_lagy1)", "log(hhi_lagy2)"))

# Clean labels
tidy_all$term <- recode(tidy_all$term,
                        "win_rate:Post" = "Win Rate × Post",
                        "log(hhi_lagy1)" = "Log HHI (t-1)",
                        "log(hhi_lagy2)" = "Log HHI (t-2)")

# Plot
ggplot(tidy_all, aes(x = term, y = estimate, color = model)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.6), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "", 
       y = "Coefficient Estimate (95% CI)",
       title = "Effect of H-1B Wins and Lagged HHI on HHI") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")
```

### Callaway Sant'anna Continuous DiD

The full Callaway, Goodman-Bacon and Sant'Anna (2024) estimator in the `contdid`
package is *not* identified in this setting. `cont_did()` compares dose-$d$ units
against untreated (dose-0) units, and every county here has `win_rate > 0` (the
minimum is 0.00285). There is no zero-dose group to recover: with the 20-petition
threshold and win rates centred near 0.32, the chance a county draws zero wins is
about $0.68^{20} \approx 0.04\%$, so a genuine zero would almost certainly be a
failed employer-name merge rather than lottery luck. Bringing in the ~1,450
never-applying counties as controls would trade the lottery randomisation for
conditional parallel trends, which is a worse design than the one we have.

What their critique *does* apply to is the linearity of `win_rate * Post`: with a
continuous dose, that single coefficient is a weighted average of causal responses
across the dose distribution. Binning the dose tests that directly, using only
variation inside the lottery-applicant sample.

```{r}
# Bin on the COUNTY-level distribution, not the observation-level one: win_rate is
# time-invariant, so cutting the full panel would just weight each county by its
# number of years.
dose_bins <- H1bdf |>
  filter(!is.na(win_rate)) |>
  distinct(COUNTY, win_rate)

dose_breaks <- quantile(dose_bins$win_rate, probs = seq(0, 1, 0.25), na.rm = TRUE)

dose_bins <- dose_bins |>
  mutate(dose_q = cut(win_rate, breaks = dose_breaks,
                      include.lowest = TRUE, labels = 1:4))

H1bdf_dose <- H1bdf |>
  filter(!is.na(win_rate)) |>
  left_join(dose_bins |> select(COUNTY, dose_q), by = "COUNTY")

# Mean dose per bin, needed to map coefficients back onto win rates
dose_summary <- dose_bins |>
  group_by(dose_q) |>
  summarise(counties = n(),
            min_dose = min(win_rate),
            mean_dose = mean(win_rate),
            max_dose = max(win_rate),
            .groups = "drop")

dose_summary
```

County fixed effects absorb `win_rate` (time-invariant) and year fixed effects
absorb `Post`, so the interactions are all that is identified. Coefficients are
therefore the post-2016 shift for each quartile *relative to quartile 1*, not an
absolute effect.

```{r}
dose_hhi <- feols(log_hhi ~ i(dose_q, Post, ref = 1) | COUNTY + YEAR,
                  data = H1bdf_dose, cluster = ~COUNTY)

dose_shannon <- feols(log_shannon ~ i(dose_q, Post, ref = 1) | COUNTY + YEAR,
                      data = H1bdf_dose, cluster = ~COUNTY)

# Linear specification for comparison
dose_linear <- feols(log_hhi ~ win_rate:Post | COUNTY + YEAR,
                     data = H1bdf_dose, cluster = ~COUNTY)


etable(dose_hhi, dose_shannon, dose_linear)
```

If the response were linear, the top quartile (mean dose 0.691) should show roughly
twice the effect of the third (mean dose 0.383). The test below reports whether it
does.

```{r}
b <- coef(dose_hhi)
V <- vcov(dose_hhi)
i3 <- "dose_q::3:Post"
i4 <- "dose_q::4:Post"

diff_34 <- b[i3] - b[i4]
se_34   <- sqrt(V[i3, i3] + V[i4, i4] - 2 * V[i3, i4])

data.frame(
  q3 = b[i3], q4 = b[i4],
  difference = diff_34, std_error = se_34,
  t_stat = diff_34 / se_34,
  p_value = 2 * pnorm(-abs(diff_34 / se_34)),
  row.names = NULL
)
```

Pre-trends, interacting event time with the dose bins. `ref2 = 1` is required:
without it the full set of `event_time` x `dose_q` dummies sums to the year
dummies and is collinear with the year fixed effects, so `fixest` silently drops
one entire quartile's path (all ten Q4 terms). Setting Q1 as the reference bin
makes the normalisation explicit and matches the levels regression above -- each
series is that quartile's path *relative to Q1*.

```{r}
dose_es <- feols(log_hhi ~ i(event_time, dose_q, ref = -1, ref2 = 1) | COUNTY + YEAR,
                 data = H1bdf_dose, cluster = ~COUNTY)

es_coef <- coef(dose_es)
es_se   <- se(dose_es)
pre_terms <- grep("^event_time::-", names(es_coef), value = TRUE)

cat("Pre-period interactions significant at 5%:",
    sum(abs(es_coef[pre_terms] / es_se[pre_terms]) > 1.96),
    "of", length(pre_terms), "\n")

length(dose_es$collin.var)  # 0 once ref2 is set
```

`iplot()` cannot draw the model above: it only handles variables built by
`i(factor, numeric)` or `i(factor)`, and `i(event_time, dose_q)` is a
factor-by-factor interaction, which raises "no variable created with i() found".
Collapsing the dose to a binary high/low split gives `iplot()` something it can
plot, at the cost of the quartile detail.

```{r}
H1bdf_dose <- H1bdf_dose |>
  mutate(high_dose = as.integer(dose_q %in% c("3", "4")))

dose_es_binary <- feols(log_hhi ~ i(event_time, high_dose, ref = -1) | COUNTY + YEAR,
                        data = H1bdf_dose, cluster = ~COUNTY)

# Same wrapper pattern as the other iplot()s in this file: base graphics, so draw
# once for the knitted output and once into the PDF written for the paper.
plot_dose_es <- function() {
  op <- par(mar = c(4.5, 4.5, 1, 1))
  on.exit(par(op))

  iplot(dose_es_binary,
        main = "",
        xlab = "Years Relative to the 2016 Lottery",
        ylab = "Effect on Log(HHI), Top Two Quartiles vs Bottom Two")
  abline(v = -0.5, lty = 3, col = "grey50")
}

plot_dose_es()

pdf("Figures/dose_event_study.pdf", width = 7, height = 5)
plot_dose_es()
dev.off()
```

The four-bin version has to be drawn by hand. Each series is one quartile's path
relative to Q1, so Q1 itself is the flat zero line by construction.

```{r}
dose_es_df <- data.frame(
  term = names(es_coef),
  estimate = as.numeric(es_coef),
  std_error = as.numeric(es_se)
) |>
  mutate(
    event_time = as.integer(sub("^event_time::(-?[0-9]+):.*$", "\\1", term)),
    dose_q = sub("^.*dose_q::([0-9]+)$", "\\1", term),
    ci_low = estimate - 1.96 * std_error,
    ci_high = estimate + 1.96 * std_error
  ) |>
  bind_rows(
    expand.grid(event_time = sort(unique(H1bdf_dose$event_time)),
                dose_q = "1") |>
      mutate(estimate = 0, std_error = 0, ci_low = 0, ci_high = 0)
  )

dose_es_plot <- ggplot(dose_es_df, aes(x = event_time, y = estimate, colour = dose_q)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_vline(xintercept = -0.5, linetype = "dotted", colour = "grey50") +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high, fill = dose_q),
              alpha = 0.12, colour = NA) +
  geom_line() +
  geom_point(size = 1.4) +
  labs(x = "Years Relative to the 2016 Lottery",
       y = "Effect on log(HHI) relative to quartile 1",
       colour = "Dose quartile", fill = "Dose quartile") +
  theme_classic(base_size = 12) +
  theme(legend.position = "bottom")

dose_es_plot

ggsave("Figures/dose_event_study_quartiles.pdf", dose_es_plot, width = 7, height = 5)
```

```{r}
# Dose-response plot: coefficients against mean dose, with quartile 1 as the
# normalised zero.
dose_plot_df <- data.frame(
  dose_q = factor(2:4),
  estimate = coef(dose_hhi),
  std_error = se(dose_hhi)
) |>
  left_join(dose_summary |> select(dose_q, mean_dose), by = "dose_q") |>
  bind_rows(data.frame(dose_q = factor(1), estimate = 0, std_error = 0,
                       mean_dose = dose_summary$mean_dose[1])) |>
  arrange(mean_dose) |>
  mutate(ci_low = estimate - 1.96 * std_error,
         ci_high = estimate + 1.96 * std_error)

dose_response <- ggplot(dose_plot_df, aes(x = mean_dose, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.015) +
  geom_point(size = 2) +
  geom_line() +
  labs(x = "Mean H-1B lottery win rate in quartile",
       y = "Effect on log(HHI) relative to quartile 1") +
  theme_classic(base_size = 12)

dose_response

ggsave("Figures/dose_response_hhi.pdf", dose_response, width = 7, height = 5)
```

### Robustness of the Dose-Response Shape

The step between the second and third quartiles is either a real non-linearity or an
artifact of how the dose is measured. Measured `win_rate` is a ratio, and its denominator
shrinks as the measured dose rises: median petitions fall monotonically across quartiles,
so the counties carrying the top of the dose distribution are exactly the ones whose win
rate is estimated from the fewest draws. Classical measurement error in a right-hand-side
variable attenuates a *linear* coefficient, but when the error is concentrated at one end
of the support it distorts the estimated *shape*, which is what a bin specification
reports.

Two checks separate the explanations. Both reuse `build_panel()` from the cap-exempt
sensitivity above, so the panel is rebuilt from source rather than filtered after the
fact.

```{r}
# Counties censored at line 110: USCIS approvals exceed LCA petitions, so the raw
# ratio is above 1. That is a merge failure, not lottery luck, and all of them sit
# in the top dose quartile -- precisely where the shape is in question.
censored_counties <- dose_raw |>
  filter(!is.na(num_petitions), num_petitions >= 20,
         initial_approval / num_petitions > 1) |>
  pull(COUNTY)

# build_panel() only checks num_petitions, so two counties with petitions recorded
# but no approval figure survive it carrying win_rate = NA. Left alone they show up
# as a silent <NA> bin in the composition table and as a "22 observations removed"
# note in every regression below. Dropped explicitly instead.
bp <- function(...) build_panel(...) |> filter(!is.na(win_rate))

panel_base <- bp("baseline")
uncensored <- setdiff(unique(panel_base$COUNTY), censored_counties)

c(censored = length(censored_counties),
  dropped_na_win_rate = n_distinct(build_panel("baseline")$COUNTY) -
                        n_distinct(panel_base$COUNTY))
```

```{r}
# Re-cutting the quartiles inside each subsample asks "is the shape still there";
# holding the baseline edges fixed asks "did these same counties move". They answer
# different questions, so both are reported.
run_dose_bins <- function(d, label, breaks = NULL) {
  cty <- d |> distinct(COUNTY, win_rate)

  if (is.null(breaks)) {
    breaks <- quantile(cty$win_rate, probs = seq(0, 1, 0.25), na.rm = TRUE)
  }

  cty <- cty |>
    mutate(dose_q = cut(win_rate, breaks = breaks,
                        include.lowest = TRUE, labels = 1:4))

  dd <- d |>
    left_join(cty |> select(COUNTY, dose_q), by = "COUNTY") |>
    filter(!is.na(dose_q))

  m <- feols(log_hhi ~ i(dose_q, Post, ref = 1) | COUNTY + YEAR,
             data = dd, cluster = ~COUNTY)

  b <- coef(m); V <- vcov(m); s <- se(m)
  i3 <- "dose_q::3:Post"; i4 <- "dose_q::4:Post"
  diff34 <- b[i3] - b[i4]
  se34   <- sqrt(V[i3, i3] + V[i4, i4] - 2 * V[i3, i4])

  tibble(
    sample      = label,
    counties    = n_distinct(dd$COUNTY),
    q2          = b["dose_q::2:Post"],
    q3          = b[i3], q3_se = s[i3],
    q4          = b[i4], q4_se = s[i4],
    q3_minus_q4 = diff34,
    p_q3_q4     = 2 * pnorm(-abs(diff34 / se34))
  )
}

base_breaks <- quantile(distinct(panel_base, COUNTY, win_rate)$win_rate,
                        probs = seq(0, 1, 0.25), na.rm = TRUE)

dose_robust <- bind_rows(
  run_dose_bins(panel_base, "baseline (>= 20 petitions)"),
  run_dose_bins(bp("baseline", min_petitions = 100),
                "restrict to >= 100 petitions"),
  run_dose_bins(bp("baseline", keep_counties = uncensored),
                "drop censored win_rate = 1"),
  run_dose_bins(bp("baseline", min_petitions = 100,
                   keep_counties = uncensored),
                "both restrictions"),
  run_dose_bins(bp("baseline", min_petitions = 100),
                ">= 100 petitions, fixed bin edges", breaks = base_breaks),
  run_dose_bins(bp("baseline", keep_counties = uncensored),
                "drop censored, fixed bin edges", breaks = base_breaks)
)

dose_robust
```

The two checks do not behave the same way. Dropping the censored counties leaves the
step intact — the third-quartile coefficient is if anything slightly larger, and the
Q3-versus-Q4 test is unchanged. Restricting to counties with at least 100 petitions
removes it: the top-quartile coefficient moves toward zero and changes sign, and with
the bin edges held fixed none of the quartile terms is distinguishable from zero. The
non-linearity therefore tracks denominator size rather than dose.

The composition table shows why the restriction bites so hard at the top.

```{r}
panel_base |>
  distinct(COUNTY, win_rate, num_petitions) |>
  mutate(dose_q = cut(win_rate, breaks = base_breaks,
                      include.lowest = TRUE, labels = 1:4)) |>
  group_by(dose_q) |>
  summarise(counties         = n(),
            median_petitions = median(num_petitions),
            pct_under_100    = round(100 * mean(num_petitions < 100)),
            surviving_at_100 = sum(num_petitions >= 100),
            .groups = "drop")
```

Because the bins are not merely noisier at the top but *built* from small-denominator
counties, the same restriction should be applied to the linear specification. It is the
headline estimate, and it is subject to the identical problem.

The floor is swept densely at the low end rather than pushed upward. Floors of 100 and
200 do answer the measurement-error question, but they select so heavily on county size
that any change is as easily read as urban heterogeneity. Floors between 1 and 40 move
the sample much less, so they separate the two explanations more cleanly.

A floor of 0 is not attainable and is not reported. 150 counties filed no petitions at
all, so `win_rate = 0/0` is undefined; `num_petitions >= 0` would admit them and
`pmin(Inf, 1)` would then score counties that filed nothing as *maximum* dose. The
lowest meaningful floor is 1.

```{r}
petition_floors <- c(1, 10, 20, 30, 40, 50, 100, 200)

# Built once and reused by the estimates, the composition table and the histograms,
# so all three describe exactly the same samples.
floor_panels <- setNames(
  lapply(petition_floors, function(f) bp("baseline", min_petitions = f)),
  petition_floors
)

# Two counties have a single reported NAICS industry, so shannon_index is exactly 0
# and log_shannon is -Inf. Both have fewer than 20 petitions, so they appear only at
# the new low floors and never touched the published sample. feols() drops them per
# model, which is correct but silent -- counted here so the drop is on the record.
sapply(floor_panels, function(d) sum(!is.finite(d$log_shannon)))

lin_floor <- function(f, dv) {
  d <- floor_panels[[as.character(f)]]
  m <- feols(as.formula(paste(dv, "~ win_rate:Post | COUNTY + YEAR")),
             data = d, cluster = ~COUNTY)
  tt <- tidy(m) |> filter(term == "win_rate:Post")

  tibble(outcome       = dv,
         min_petitions = f,
         counties      = n_distinct(d$COUNTY),
         estimate      = tt$estimate,
         std.error     = tt$std.error,
         p.value       = tt$p.value)
}

threshold_table <- bind_rows(
  lapply(petition_floors, lin_floor, dv = "log_hhi"),
  lapply(petition_floors, lin_floor, dv = "log_shannon")
)

threshold_table
```

The estimate is not monotone in the floor. It peaks at 20 and decays in *both*
directions: for `log_hhi` it runs +0.033 at a floor of 1, -0.051 at 10, -0.064 at 20,
-0.058 at 30, -0.013 at 40 and -0.004 at 50, and `log_shannon` mirrors it with the
opposite sign throughout. Neither candidate explanation predicts that shape on its own.
Classical measurement error attenuates toward zero, so cleaning the dose by raising the
floor should *strengthen* the estimate; instead it destroys it. Urban heterogeneity
predicts a monotone drift as the sample shifts, not a peak.

That the maximum sits exactly at the published floor of 20 deserves to be stated
plainly. The threshold was chosen ex ante, for the stated reason that win rates built on
few petitions are mechanically noisy, and not tuned against the result. But the estimate
is clearly sensitive to it, and a reader is entitled to see the whole curve rather than
one point on it.

The sign flip at a floor of 1 is not a competing result. That sample is not "unselected",
it is maximally noisy: median petitions of 6, a quarter of counties censored at
`win_rate = 1`, and the dose for a county with one or two petitions is close to a coin
flip. Both ends of the sweep are contaminated, in opposite ways.

```{r}
# ESTAB is total establishments in the county, so it stands in for county size and
# shows directly how much of each floor's movement is compositional.
floor_composition <- bind_rows(lapply(petition_floors, function(f) {
  d   <- floor_panels[[as.character(f)]]
  cty <- d |> distinct(COUNTY, win_rate, num_petitions)
  est <- d |> filter(YEAR == 2016) |> distinct(COUNTY, ESTAB)

  tibble(min_petitions = f,
         counties      = nrow(cty),
         med_petitions = median(cty$num_petitions),
         med_estab     = median(est$ESTAB),
         mean_win_rate = mean(cty$win_rate),
         sd_win_rate   = sd(cty$win_rate),
         pct_zero      = 100 * mean(cty$win_rate == 0),
         pct_censored  = 100 * mean(cty$win_rate == 1))
}))

floor_composition
```

The composition table both confirms the concern about high floors and limits how much
work it can do. Median establishments per county rise from about 2,000 at a floor of 1 to
9,300 at 20 and 19,600 at 200, so floors of 100 and 200 are indeed heavily urban-selected.
But the estimate has already collapsed by a floor of 40, where median establishments are
12,600 — a 36% shift from the floor-of-20 sample, against the 4.7-fold shift between
floors 1 and 20. The result therefore dies across a range where the sample composition
barely moves, which is difficult to attribute to urban selection alone.

The dose distribution deteriorates at the bottom rather than the top: the share of
counties censored at `win_rate = 1` runs 27% at a floor of 1, 8% at 10, 4% at 20 and
under 2% by 40, and counties at `win_rate = 0` disappear entirely by a floor of 30.

```{r}
# Distributions are county-level, not county-year. win_rate is time-invariant, so
# histogramming the panel would weight each county by its number of years.
dist_floors <- c(1, 10, 20, 30, 40)

winrate_by_floor <- bind_rows(lapply(dist_floors, function(f) {
  floor_panels[[as.character(f)]] |>
    distinct(COUNTY, win_rate) |>
    mutate(floor = factor(paste0(">= ", f, " petitions"),
                          levels = paste0(">= ", dist_floors, " petitions")))
}))

winrate_floor_plot <- ggplot(winrate_by_floor, aes(x = win_rate)) +
  geom_histogram(bins = 30, fill = "lightgreen", colour = "black") +
  facet_wrap(~ floor, ncol = 3, scales = "free_y") +
  labs(x = "Win Rate", y = "Count") +
  theme_classic(base_size = 12)

winrate_floor_plot

ggsave("Figures/winrate_distribution_by_floor.pdf", winrate_floor_plot,
       width = 9, height = 6)
```

The histograms show what the summary statistics compress. At a floor of 1 the
distribution is barely a distribution: a spike at 0, a large spike at 1, and mass at the
simple fractions a handful of petitions can produce. Raising the floor pulls it toward a
single mode near 0.33 and strips the spikes at both ends. The dose only starts behaving
like a continuous measure somewhere around a floor of 20 to 30 — which is also, awkwardly,
where the estimated effect is disappearing.

One caveat still governs the section. Restricting on `num_petitions` selects on county
size, so measurement error and population composition can never be fully separated by
cutting the sample, only bounded. The floor sweep narrows the gap considerably by showing
the collapse happens where composition moves least, but settling it requires weighting by
the denominator or an errors-in-variables correction that uses petition counts to model
the noise directly.

```{r}
threshold_plot <- ggplot(threshold_table,
                         aes(x = factor(min_petitions), y = estimate,
                             colour = outcome, group = outcome)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error,
                    ymax = estimate + 1.96 * std.error),
                width = 0.08, position = position_dodge(width = 0.3)) +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  labs(x = "Minimum petitions required for inclusion",
       y = "Estimated win rate x Post coefficient",
       colour = "Outcome") +
  theme_classic(base_size = 12) +
  theme(legend.position = "bottom")

threshold_plot

ggsave("Figures/robustness_petition_threshold.pdf", threshold_plot,
       width = 7, height = 5)
```

## DiD Using Driscoll-Kray Standard Errors

### Original DiD using HHI

```{r}
did_countyDK <- feols(log_hhi ~ win_rate * Post | COUNTY + YEAR, data = H1bdf, panel.id = ~COUNTY + YEAR, se = "dk")

event_studydk <- feols(log_hhi ~ i(event_time, win_rate, ref = -1) + factor(COUNTY) + factor(YEAR), data = H1bdf, panel.id = ~COUNTY + YEAR, se = "dk")

summary(did_countyDK)

plot_event_study_dk <- function() {
  op <- par(mar = c(4.5, 4.5, 1, 1))
  on.exit(par(op))

  iplot(event_studydk,
        main = "",
        xlab = "Years Relative to Event",
        ylab = "Effect of Win Rate on Log(Counts)")
}

plot_event_study_dk()

pdf("Figures/event_study_dk.pdf", width = 7, height = 5)
plot_event_study_dk()
dev.off()
```

### Placebo DiD using HHI

```{r}
did_countyDK_Placebo <- feols(log_hhi ~ win_rate * Post_Placebo + i(COUNTY) + i(YEAR), data = H1bdf, panel.id = ~COUNTY + YEAR, se = "dk")

summary(did_countyDK_Placebo)
```

Using a county clustering standard error treatment, the P-value for the original regression is marginally statistically significant (P \< 0.1) but the P-value for the placebo is not statistically significant. While using Driscoll-Kraay SEs, the original regression is highly statistically significant (P \< 0.01) while the placebo is only statistically significant at the 5% level (P \< 0.05). Since both standard error treatments show that the placebo P-value is larger than the original P-value, this supports the parallel trends assumption and strenghthens the causal interpretation. However, using Driscoll-Kraay SEs seems like it is overinflating P values across the board, because all P values including in all the DiD regressions and event study are statistically significant to a very small level.

### Conducting regression using Shannon Diversity Index

```{r}
did_continuous_shannon <- feols(log_shannon ~ win_rate * Post | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

summary(did_continuous_shannon)
```

```{r}
placebo_model_shannon <- feols(log_shannon ~ win_rate * Post_Placebo | COUNTY + YEAR, data = H1bdf, cluster = "COUNTY")

summary(placebo_model_shannon)
```

### Excluding 2020 and 2021

```{r}
did_continuous_hhi_subset <- feols(log_hhi ~ win_rate * Post + i(COUNTY) + i(YEAR), data = H1bdf %>% filter(YEAR < 2020), cluster = "COUNTY")

summary(did_continuous_hhi_subset)

haven::write_dta("H1bdf.dta", data = H1bdf)
```

Excluding 2020 and 2021 yields a similar coefficient to the main regression. Therefore, this suggests my coefficient is likely to be causal.

## Extra Visualizations

```{r}
library(tidycensus)
library(ggplot2)
library(sf)


# Download county boundaries for all states
us_counties_2022 <- tigris::counties(cb = TRUE, 
                     resolution = "20m", 
                     progress_bar = FALSE,
                     year = 2022) |>
  filter(NAME != "Puerto Rico") |>
  tigris::shift_geometry()

# Convert FIPS to match your dataset (strip leading zeros)
us_counties_2022 <- us_counties_2022 %>%
  mutate(fips = as.character(as.numeric(GEOID)))

# Check structure
head(us_counties_2022)

map_data <- left_join(us_counties_2022, CBP_diversity, by = c("AFFGEOID" = "GEO_ID"))
```

```{r}
# Calculate percentiles
lower <- quantile(map_data$shannon_index, 0.01, na.rm = TRUE)
upper <- quantile(map_data$shannon_index, 0.99, na.rm = TRUE)

# Winsorize
map_data <- map_data %>%
  mutate(residual_naics = pmin(pmax(log_shannon, lower), upper))

# Percentile
map_data <- map_data |>
  mutate(HHI_pctile = percent_rank(hhi)*100)
```

```{r}
library(viridis)

# Reproject to a U.S.-focused CRS
map_data <- st_transform(map_data, crs = 5070)  # USA Albers Equal-Area

hhi_map <- ggplot(map_data) +
  geom_sf(aes(fill = HHI_pctile), color = "white", lwd = 0.05) +
  scale_fill_viridis_c(option = "plasma", direction = -1, name = "HHI Percentile") +
  labs(caption = "Source: CBP & Census Data") +
  theme_minimal() +
  theme(
    legend.position = c(0.9, 0.15),  # Moves legend inside map
    legend.background = element_rect(fill = "white", color = "black"),
    legend.key = element_rect(fill = "white")
  ) +
  coord_sf(
    xlim = c(-2500000, 2500000),  # Manually set U.S. bounding box in meters
    ylim = c(-70000, 3200000),  
    expand = FALSE
  )

hhi_map

ggsave("Figures/HHI_map.pdf", hhi_map, width = 12, height = 8, bg = "white")
```

```{r}
# Table 2 data
event_time <- c(-5, -4, -3, -2, 0, 1, 2, 3, 4, 5)
coeff <- c(-0.003306, -0.003757, 0.004117, 0.003612, -0.067812,
          -0.066438, -0.065368, -0.064380, -0.057434, -0.058553)
se <- c(0.01745044, 0.01295252, 0.00914417, 0.00636837, 0.04351837,
        0.04038470, 0.03893465, 0.03986916, 0.03947110, 0.03828369)

# 90% confidence intervals
upper <- coeff + 1.645 * se
lower <- coeff - 1.645 * se

# Data frame
df <- data.frame(event_time, coeff, upper, lower)

# Add baseline (year -1) manually
df <- rbind(df, data.frame(
  event_time = -1,
  coeff = 0,
  upper = NA,
  lower = NA
))

# Sort the data so -1 is in the right place
df <- df[order(df$event_time), ]

# Plot
event_study_hhi_ribbon <- ggplot(df, aes(x = event_time, y = coeff)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  labs(title = "Event Study: Effect of H-1B Lottery Win Rate Over Time",
       x = "Years Since Lottery (t)",
       y = "Effect of Win Rate on log(HHI)") +
  theme_minimal()

event_study_hhi_ribbon

ggsave("Figures/event_study_hhi_ribbon.pdf", event_study_hhi_ribbon, width = 7, height = 5)

```

### Number of days until H-1B cap is filled

```{r}
years <- 2002:2020

days_open <- c(365, 365, 322, 183, 131, 55, 2, 5, 264, 300, 235, 71, 5, 5, 5, 5, 5, 5, 5)

cap_filled <- data.frame(FiscalYear = years, DaysOpen = days_open)

h1b_cap_days <- ggplot(cap_filled, aes(x = FiscalYear, y = DaysOpen)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2002, 2020, by = 2)) +
  scale_y_continuous(breaks = seq(0,360, by = 60)) +
  theme_classic() +
  labs(
    x = "Fiscal Year",
    y = "H-1B Cap-Subject Application Period\n(Days from Start of Filing to Final Receipt Date)"
  )

h1b_cap_days

ggsave("Figures/h1b_cap_days.pdf", h1b_cap_days, width = 7, height = 5)

```

### Event Study Using DK SEs

```{r}
event_study_hhidk <- feols(log_hhi ~ i(event_time, win_rate, ref = -1) + factor(COUNTY) + factor(YEAR), data = H1bdf, panel.id = ~COUNTY + YEAR, se = "dk")

summary(event_study_hhidk)
```

```{r}
# Create the data frame (copied from the previous step)
dk_event_study <- data.frame(
  event_time = c(-5, -4, -3, -2, 0, 1, 2, 3, 4, 5),
  estimate = c(
    -0.003306, -0.003757, 0.004117, 0.003612,
    -0.067812, -0.066438, -0.065368, -0.064380,
    -0.057434, -0.058553  # Add NA placeholder if t=5 missing
  ),
  std_error = c(
    rep(0.000010, 10)
  )
)

# Add event_time = -1 with estimate and std_error set to 0
dk_event_study <- bind_rows(
  dk_event_study,
  data.frame(event_time = -1, estimate = 0, std_error = 0)
) %>%
  arrange(event_time) %>%
  mutate(
    lower = estimate - 1.645 * std_error,
    upper = estimate + 1.645 * std_error
  )

# Plot
event_study_dk_ribbon <- ggplot(dk_event_study, aes(x = event_time, y = estimate)) +
  geom_line(color = "black") +
  geom_point(color = "black") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.3, fill = "gray") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(
    title = "Event Study with Driscoll-Kraay SEs (HHI)",
    x = "Years Since Lottery (t)",
    y = "Effect of Win Rate on log(HHI)"
  ) +
  theme_minimal()

event_study_dk_ribbon

ggsave("Figures/event_study_dk_ribbon.pdf", event_study_dk_ribbon, width = 7, height = 5)


```
