---
title: "ECON 4198W Data Cleaning Part 2"
author: "Spencer Ma"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 7)
knitr::opts_chunk$set(warning = FALSE, results = 'hide', message = F)
library(here)

# Pin the project root to this file's location rather than letting here() guess
# from the session's working directory. Every Excel path below, and the Figures
# export, depend on this resolving correctly.
here::i_am("H-1B_Data_Cleaning.Rmd")

# Run log. Appends to logs/H-1B_Data_Cleaning.log; see log_setup.R.
source(here("log_setup.R"))
log_init("H-1B_Data_Cleaning")
```

# Data Cleaning H-1B Data

## Uploading the Data

```{r}
library(data.table)
library(readxl)

# Define the file path
file_path <- here("Datasets", "H-1B Data",  "H-1B_Disclosure_Data_FY16.xlsx")

# Read the sheet into a data.table
H1B2016_Disclosure_Data <- as.data.table(read_excel(file_path, sheet = 1))
log_dim("LCA disclosure FY16", H1B2016_Disclosure_Data)
```

```{r}
H1BFY2017_employer_data_hub <- as.data.table(
  read_excel(here("Datasets", "H-1B Data", "H-1B Employer Data Hub (2017).xlsx"), sheet = 1)
)
log_dim("USCIS Data Hub FY17", H1BFY2017_employer_data_hub)
```

## Visualization of the 2016 H-1B Disclosure Data

```{r}
library(dplyr)
library(lubridate)

H1b2016Viz <- H1B2016_Disclosure_Data[CASE_SUBMITTED >= "2016-01-01" & CASE_SUBMITTED <= "2016-12-31"]

# Create a 'week' column
H1b2016Viz[, week := floor_date(CASE_SUBMITTED, unit = "week", week_start = 1)]

# Count number of cases per week
weekly_counts <- H1b2016Viz[, .N, by = week]

# Convert week column to Date format
weekly_counts[, week := as.Date(week)]

# Remove December Entry
weekly_counts <- weekly_counts %>% filter(week != "2015-12-28")

# View data structure
head(weekly_counts)
```

### Visualizing other years

All eight appendix panels are built by the loop below. It handles one year at a
time and never holds two disclosure files at once, which is the equivalent of
Stata's `use ..., clear`. The mechanism is scope rather than an explicit command:
everything `make_h1b_graph()` creates is local to that call and is released when
it returns, so nothing accumulates across iterations.

Three wrinkles this has to work around:

-   DOL renamed the submission-date column in FY2015. The 2012-2014 files call it
    `LCA_CASE_SUBMIT`; 2015 onward call it `CASE_SUBMITTED`.
-   The files are wide and slow to read (FY2019 has 260 columns), so
    `read_submit_dates()` tells readxl to skip every column except the date. That
    is what keeps this practical to run in one go.
-   Reading is still the expensive step, so a year whose PDF already exists is
    skipped. Delete the PDF to force a rebuild.

```{r}
library(ggplot2)

h1b_files <- c(
  "2012" = "LCA_FY2012_Q4.xlsx",
  "2013" = "LCA_FY2013.xlsx",
  "2014" = "H-1B_FY14_Q4.xlsx",
  "2015" = "H-1B_Disclosure_Data_FY15_Q4.xlsx",
  "2016" = "H-1B_Disclosure_Data_FY16.xlsx",
  "2017" = "H-1B_Disclosure_Data_FY17.xlsx",
  "2018" = "H-1B_Disclosure_Data_FY2018_EOY.xlsx",
  "2019" = "H-1B_Disclosure_Data_FY2019.xlsx"
)

# Read the submission-date column and nothing else.
read_submit_dates <- function(path) {
  nms <- names(read_excel(path, sheet = 1, n_max = 0))

  date_col <- intersect(c("CASE_SUBMITTED", "LCA_CASE_SUBMIT"), nms)[1]
  if (is.na(date_col)) stop("No submission-date column in ", basename(path))

  col_types <- rep("skip", length(nms))
  col_types[which(nms == date_col)[1]] <- "date"

  dt <- as.data.table(read_excel(path, sheet = 1, col_types = col_types))
  setnames(dt, 1, "submitted")
  dt[, submitted := as.Date(submitted)]
  dt[]
}

# One year in, one PDF out. Nothing created in here outlives the call.
make_h1b_graph <- function(year, file) {
  submissions <- read_submit_dates(here("Datasets", "H-1B Data", file))

  in_year <- submissions[submitted >= as.Date(paste0(year, "-01-01")) &
                         submitted <= as.Date(paste0(year, "-12-31"))]

  rm(submissions)  # drop the full-file table before doing anything else
  gc()

  in_year[, week := floor_date(submitted, unit = "week", week_start = 1)]
  weekly_counts <- in_year[, .N, by = week][order(week)]

  # Drop the partial leading week that starts in the previous December
  weekly_counts <- weekly_counts[week >= as.Date(paste0(year, "-01-01"))]

  weekly_plot <- ggplot(weekly_counts, aes(x = week, y = N)) +
    geom_line(color = "purple", size = 1) +
    geom_point(size = 1) +
    labs(title = paste("Weekly H-1B Cases Submitted in", year),
         x = "Week",
         y = "Number of Cases Submitted") +
    scale_y_continuous(breaks = seq(0, 55000, by = 5000)) +
    scale_x_date(date_labels = "%b %d", date_breaks = "1 month") +
    theme_minimal(base_size = 16)

  ggsave(here("Figures", paste0("H-1B_graph_", year, ".pdf")),
         weekly_plot, width = 12, height = 8, bg = "white")
}

# Loop variable is viz_year, not year: lubridate::year() is used further down
# this file and a global named `year` is a trap worth not setting.
for (viz_year in names(h1b_files)) {
  out_file <- here("Figures", paste0("H-1B_graph_", viz_year, ".pdf"))

  if (file.exists(out_file)) {
    message("Skipping ", viz_year, ": ", basename(out_file), " already exists")
    next
  }

  message("Building ", viz_year, " from ", h1b_files[[viz_year]])
  make_h1b_graph(viz_year, h1b_files[[viz_year]])
  gc()
}
```

# Cleaning the Disclosure Dataset

```{r}
Disclosure_data <- H1B2016_Disclosure_Data[VISA_CLASS == "H-1B"]

Disclosure_data <- Disclosure_data[!(CASE_STATUS %in% c("WITHDRAWN", "CERTIFIED-WITHDRAWN", "DENIED"))]

#Convert CASE_SUBMITTED to Date format
Disclosure_data <- Disclosure_data[, CASE_SUBMITTED := as.Date(CASE_SUBMITTED)]

Likely_lottery <- Disclosure_data[CASE_SUBMITTED >= as.Date("2016-03-01") & CASE_SUBMITTED < as.Date("2016-04-03")]
```

I removed cases with non H-1B visas such as the E-3 Australian, the Singapore H-1B1 and the Chile H-1B1. I also excluded cases where the LCA was withdrawn, certified, withdrawn, or denied because those cases never entered the H-1B lottery.

As described in Mahajan et. al, LCA cases submitted from March to April are more likely to be lottery-bound. That is why I removed any LCAs not submitted between March 1st and April 5th, which was the time USCIS started accepting H-1B petitions.

## Removing Cap-Exempt Employers from Both Datasets

```{r}
# Three cap-exempt filters rather than one, so the downstream analysis can report
# how sensitive the win rate is to where this line is drawn. The LCA data has no
# cap-exempt flag, so any keyword rule is a guess; the point of "none" and "broad"
# is to bracket that guess, not to be right.
#
#   none     - no exclusion at all. Upper bound on the denominator.
#   baseline - the filter used in the published paper. Unchanged.
#   broad    - deliberately over-inclusive. Catches genuine cap-exempt employers
#              but also plain for-profits ("trust" hits SunTrust, "association"
#              hits trade groups). Lower bound on the denominator, NOT an estimate.
cap_exempt_specs <- list(
  none = character(0),

  baseline = c("university", "univ", "college", "school", "research", "hospital",
               "medical center", "foundation", "institute", "government",
               "public health"),

  broad = c("university", "univ", "college", "school", "research", "hospital",
            "medical center", "foundation", "institute", "government",
            "public health", "medical", "health", "clinic", "academy",
            "nonprofit", "non-profit", "charit", "trust", "museum", "library",
            "seminary", "regents", "trustees", "board of ", "district",
            "city of", "county of", "state of", "department of", "authority",
            "center for", "society", "association", "institution", "laborator")
)

# Tag rather than drop. Every spec's totals are then a different weighted sum over
# the same rows, so the pipeline runs once instead of three times -- which matters
# because re-running it means re-reading the 75-270MB disclosure files.
flag_keep <- function(employer_name, keywords) {
  if (length(keywords) == 0) return(rep(TRUE, length(employer_name)))
  !grepl(paste(keywords, collapse = "|"), tolower(employer_name))
}

Lottery_winners <- copy(H1BFY2017_employer_data_hub)

for (spec in names(cap_exempt_specs)) {
  kw <- cap_exempt_specs[[spec]]
  Likely_lottery[,  (paste0("keep_", spec)) := flag_keep(EMPLOYER_NAME, kw)]
  Lottery_winners[, (paste0("keep_", spec)) := flag_keep(`Employer (Petitioner) Name`, kw)]
}

# How much each spec removes, in employers (not yet in petitions).
sapply(names(cap_exempt_specs), function(s)
  c(LCA    = mean(!Likely_lottery[[paste0("keep_", s)]]),
    USCIS  = mean(!Lottery_winners[[paste0("keep_", s)]])))
```

I removed keywords pertaining to cap-exempt employers, including universities, research institutions, and nonprofits.

## Extra Cleaning

```{r}
library(tidyverse)
library(janitor)

# Convert column names to lowercase and clean up spaces
Likely_lottery <- Likely_lottery %>%
  clean_names()  # Converts column names to snake_case

# Convert date columns to proper date format
Likely_lottery <- Likely_lottery %>%
  mutate(
    case_submitted = ymd(case_submitted),
    decision_date = ymd(decision_date),
    employment_start_date = ymd(employment_start_date),
    employment_end_date = ymd(employment_end_date)
  )

# Standardize employer names (lowercase, trim spaces)
Likely_lottery <- Likely_lottery %>%
  mutate(employer_name = tolower(trimws(employer_name)))

# Extract employment year for merging
Likely_lottery <- Likely_lottery %>%
  mutate(employment_year = year(employment_start_date))

```

```{r}
# Clean column names
Lottery_winners <- Lottery_winners %>%
  clean_names()

# Ensure Fiscal Year is numeric
Lottery_winners <- Lottery_winners %>%
  mutate(fiscal_year = as.numeric(fiscal_year))

# Standardize employer names (lowercase, trim spaces)
Lottery_winners <- Lottery_winners %>%
  mutate(employer_petitioner_name = tolower(trimws(employer_petitioner_name)))

# Convert Petitioner Zip Code to string (avoid numeric formatting issues)
Lottery_winners <- Lottery_winners %>%
  mutate(petitioner_zip_code = as.character(as.integer(petitioner_zip_code)))

# Drop rows where employer name is missing
Lottery_winners <- Lottery_winners %>%
  filter(!is.na(employer_petitioner_name))

#write.csv(Likely_lottery, "likely_lottery.csv")
#write.csv(Lottery_winners, "lottery_winners.csv")
```

# Aggregating by county

```{r}
library(tidycensus)
library(tigris)

zip_county <- read_excel("ZIP_COUNTY_122016.xlsx")

# Ensure ZIP codes are character for merging
zip_county$ZIP <- as.character(zip_county$ZIP)

Lottery_winners$petitioner_zip_code <- as.character(Lottery_winners$petitioner_zip_code)

# Merge lottery winners dataset with ZIP-to-County crosswalk
uscis_with_county <- Lottery_winners %>%
  left_join(zip_county, by = c("petitioner_zip_code" = "ZIP"))

# Load Census FIPS-to-County mapping
data(fips_codes)  # This dataset comes with tidycensus

# Extract relevant columns (state, county, FIPS code)
county_fips <- fips_codes %>%
  select(state_code, county_code, county)

# Ensure COUNTY column is character for merging
uscis_with_county$COUNTY <- as.character(uscis_with_county$COUNTY)

# Merge to get actual county names
uscis_with_county <- uscis_with_county %>%
  left_join(county_fips, by = c("COUNTY" = "county_code"))
```

## Handling unmatched ZIP codes

```{r}
unmatched_zips <- Lottery_winners %>%
  anti_join(zip_county, by = c("petitioner_zip_code" = "ZIP"))

# Count how many ZIPs are unmatched
nrow(unmatched_zips)

# View a sample of unmatched ZIP codes
head(unmatched_zips$petitioner_zip_code)

# Ensure ZIP codes are character type
Lottery_winners$petitioner_zip_code <- as.character(Lottery_winners$petitioner_zip_code)
zip_county$ZIP <- as.character(zip_county$ZIP)

# Pad missing leading zeros (ensures all ZIPs are 5 digits)
Lottery_winners$petitioner_zip_code <- str_pad(Lottery_winners$petitioner_zip_code, width = 5, pad = "0")
zip_county$ZIP <- str_pad(zip_county$ZIP, width = 5, pad = "0")

uscis_with_county <- Lottery_winners %>%
  left_join(zip_county, by = c("petitioner_zip_code" = "ZIP"))

uscis_with_county <- uscis_with_county %>%
  filter(!industry_naics_code %in% c("611310", "622110", "541711", "541712", "921110"))

```

```{r}
unmatched_final <- uscis_with_county %>%
  filter(is.na(COUNTY))

# View a sample of remaining unmatched ZIPs
head(unmatched_final$petitioner_zip_code)

unique(unmatched_final$petitioner_zip_code)

#write.csv(uscis_with_county, "uscis_with_county.csv")
```

## Aggregating USCIS data by county

```{r}
# Keep only records with initial approvals and exclude continuing approvals
uscis_filtered <- uscis_with_county %>%
  filter(as.numeric(initial_approval) > 0)

# Aggregate the dataset at the county level, summing only initial_approval.
# Multiplying by the logical keep_* flag zeroes out excluded employers, so the
# three cap-exempt specs are three sums over one pass of the same rows.
county_aggregated <- uscis_filtered %>%
  group_by(COUNTY) %>%
  summarise(
    initial_approval_none     = sum(as.numeric(initial_approval) * keep_none,     na.rm = TRUE),
    initial_approval_baseline = sum(as.numeric(initial_approval) * keep_baseline, na.rm = TRUE),
    initial_approval_broad    = sum(as.numeric(initial_approval) * keep_broad,    na.rm = TRUE),
    RES_RATIO = mean(as.numeric(RES_RATIO), na.rm = TRUE),  # Taking mean since it's a ratio
    BUS_RATIO = mean(as.numeric(BUS_RATIO), na.rm = TRUE),
    OTH_RATIO = mean(as.numeric(OTH_RATIO), na.rm = TRUE),
    TOT_RATIO = mean(as.numeric(TOT_RATIO), na.rm = TRUE)
  )

# View the first few rows of the aggregated dataset
head(county_aggregated)

# National approval totals under each spec.
colSums(county_aggregated[, grep("^initial_approval_", names(county_aggregated))])
```

## Aggregating LCA data at the county level

```{r}
library(stringr)

# Convert ZIP codes to character and pad leading zeros if necessary
lca_data <- Likely_lottery %>%
  mutate(employer_postal_code = str_pad(as.character(employer_postal_code), width = 5, pad = "0"))

zip_county <- zip_county %>%
  mutate(ZIP = str_pad(as.character(ZIP), width = 5, pad = "0"))

# Merge LCA dataset with ZIP-to-County crosswalk
lca_with_county <- lca_data %>%
  left_join(zip_county, by = c("employer_postal_code" = "ZIP"))

sum(is.na(lca_with_county$COUNTY))

```

### Debugging Fips Codes

```{r}
# Find ZIP codes that did not match
unmatched_zips <- lca_data %>%
  anti_join(zip_county, by = c("employer_postal_code" = "ZIP"))

# Print unique unmatched ZIP codes
#print(unique(unmatched_zips$worksite_postal_code))

# Extract only the first 5 digits if ZIP codes have ZIP+4 formatting
lca_data <- lca_data %>%
  mutate(worksite_postal_code = str_sub(employer_postal_code, 1, 5))

zip_county <- zip_county %>%
  mutate(ZIP = str_sub(ZIP, 1, 5))

lca_with_county <- lca_data %>%
  left_join(zip_county, by = c("employer_postal_code" = "ZIP"))

sum(is.na(lca_with_county$COUNTY))  # Check missing counties again

```

```{r}
# Find unmatched ZIPs after merging
unmatched_zips <- lca_data %>%
  anti_join(zip_county, by = c("employer_postal_code" = "ZIP"))

# Print unique unmatched ZIPs
#print(unique(unmatched_zips$worksite_postal_code))

# Remove non-numeric characters and trim spaces
lca_data <- lca_data %>%
  mutate(employer_postal_code = str_trim(employer_postal_code),  # Remove spaces
         employer_postal_code = str_extract(employer_postal_code, "\\d+"),  # Keep only numbers
         employer_postal_code = str_pad(employer_postal_code, width = 5, pad = "0"))  # Ensure 5 digits

lca_with_county <- lca_data %>%
  left_join(zip_county, by = c("employer_postal_code" = "ZIP"))

sum(is.na(lca_with_county$COUNTY))  # Check missing counties again
```

### Labeling missing counties as unknown

```{r}
lca_with_county <- lca_with_county %>%
  mutate(COUNTY = ifelse(is.na(COUNTY), "Unknown", COUNTY))

table(is.na(lca_with_county$COUNTY))
```

### Aggregating LCA data

```{r}
# Aggregate by the number of workers in each county
county_aggregated_lca <- lca_with_county %>%
  filter(!is.na(COUNTY)) %>%  # Remove missing county entries
  group_by(COUNTY) %>%  
  summarise(  # Sum total workers per county, once per cap-exempt spec
    num_petitions_none     = sum(total_workers * keep_none,     na.rm = TRUE),
    num_petitions_baseline = sum(total_workers * keep_baseline, na.rm = TRUE),
    num_petitions_broad    = sum(total_workers * keep_broad,    na.rm = TRUE),
    .groups = "drop"
  )

# View aggregated dataset
head(county_aggregated_lca)

#write.csv(county_aggregated, "USCIS_county_aggregated.csv")
#write.csv(county_aggregated_lca, "LCA_county_aggregated.csv")
```

# Final Merge

```{r}
Lottery_Final <- county_aggregated |>
  full_join(county_aggregated_lca, by = "COUNTY")

Lottery_Final <- Lottery_Final |>
  select(-RES_RATIO, -BUS_RATIO, -OTH_RATIO, -TOT_RATIO)

head(Lottery_Final)

Lottery_Final <- Lottery_Final |>
  mutate(across(c(starts_with("initial_approval_"), starts_with("num_petitions_")),
                as.integer))

# `initial_approval` and `num_petitions` are kept as aliases for the baseline spec
# so H-1B_Data.qmd reproduces the published results with no change. The _none and
# _broad columns are additive; the sensitivity section reads them by name.
Lottery_Final <- Lottery_Final |>
  mutate(
    initial_approval = initial_approval_baseline,
    num_petitions    = num_petitions_baseline
  )

write.csv(Lottery_Final, "Lottery_Applications_vs_Approvals.csv", row.names = FALSE)
log_dim("Lottery_Final", Lottery_Final)
log_write("Lottery_Applications_vs_Approvals.csv")
log_done()
```
